---
title: "Project 1"
author: "Ama Agyemang-Duah"
date: 01/03/2025
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

# Conference Survey Project

This project is an analysis of data gathered from a survey of participants at a conference in
the Netherlands.

### IMPPORTING OF PACKAGES REQUIRED FOR THE ANALYSIS & LOADING SURVEY DATA

```{r message=FALSE, warning=FALSE}

library(readr) #importing package to read CSV file(*in tidyverse)
library(visdat) #this package makes it easy to visualise the whole dataset
library(dplyr) # importing this package to access data manipulating functions (*in tidyverse)
library(stringr) #this package provides string functions(*in tidyverse)
library(janitor) #importing this package to clean dataset
library(tidyr) #importing make data consistent form. (*in tidyverse)
library(ggplot2) #importing package to plot graphs and plots (*in tidyverse)

survey <- read_csv("data/survey.csv/survey.csv") #importing survey data 


```

### PART 1: EXPLORING AND CLEANING SURVEY DATA

```{r}

dim(survey)  #retrieving dimensions of survey data to know number of participants and number of columns

nrow(survey) #another way to retrieve the number of participants

names(survey) <-str_to_lower(str_replace_all(names(survey)," ","_")) #replaced all spaces with _ and changed the column names to lower case

names(survey) #displays column names

survey <- survey %>%
  rename("no" = `...1`) #renamed first column to `no`

survey %>% clean_names() #cleaning data. function is from janitor

vis_miss(survey) #This is to check whether there are any miss in values or observations

```

COMMENT ON ORIGINAL DATA: THERE ARE NO MISSING VALUES

#### -\>Showing The Number Of Survey Participants, Data types of attibutes or columnts & Top 3 Universities

```{r}

str(survey) #shows the structure of the data as well as data types

#Storing the number of participants per university in descending order in grouped_uni_data table
grouped_uni_data <-survey %>%
  group_by(university) %>%
  summarise(count=n())  %>%
  arrange(desc(count))

#Showing the top 3 of the list above 
head(grouped_uni_data,3)
```

#### -\>Survey Summary Findings

THERE ARE:

1.  30 Survey Participants

2.  a)no, number_of_employees,
    number_of_phd_students,number_of_postdocs, number_of_professors are
    double data types

    b)university, position,sizes are character data types

3.  The top 3 Universities with the most survey participants:

    Deft University of Technology,

    Eindhoven University of Technology,

    University of Groningen

### PART 2: INFORMATION ON THE NUMBER OF SURVEY PARTICIPANTS FOR DIFFERENT POSITIONS

```{r}

#group survey data by position and count the number of participants in each position
grouped_position_data <- survey %>%
        group_by(position)%>%
        summarise(count = n())

View(grouped_position_data) #Viewing the count of groupes of positions

```

#### -\>Bar Chart to show the Number Of Survey Participants For Different Positions

```{r}
      #plotting bar chart of the total count of  per positions
      ggplot(data = grouped_position_data,
       mapping = aes(x = count, 
                     y = position, 
                    fill = position)) + 
      geom_col() +
      labs(title = "Number of survey participant by position",
       subtitle = "The most common positions were Postdocs and PhDs",
       caption = "Source: Survey Dataset",
       x = "count",
       y = "position")
      
```

### PART 3: RECREATING THE `Sizes` COLUMN CORRECTLY BASED ON NUMBER OF EMPLOYEES

```{r}
  #using case_when to create corrected_sizes column based on new conditions
  #small (0 to 20), "Medium (21 to 50), "Large (51 to 200), "Extra Large (over 200)"
  corrected_survey_data <-survey %>%
  mutate (corrected_sizes = case_when(
           (`number_of_employees`) <= 20 ~ "Small (0 to 20)",
          (`number_of_employees`) > 20 & (`number_of_employees`) <= 50 ~ "Medium (21 to 50)",
          (`number_of_employees`) > 51 & (`number_of_employees`) <= 200 ~ "Large (51 to 200)",
          TRUE ~ "Extra Large (over 200)"))

  #Can also use cut functions eg_cut(data, breaks=c(0,20,50,200,Inf),c( "Small (0 to 20)","Medium (21 to 50)","Large (51 to 200)","Extra Large (over 200)"))

names(corrected_survey_data) #viewing all column names including newly added corrected_sizes
```

#### -\>Retrieving the Count of each Corrected Size Category

```{r}

  #counting the number of participants in each corrected_sizes category
  grouped_corrected_sizes_data <- corrected_survey_data %>%
        group_by(corrected_sizes)%>%
        summarise(count = n())

  View(grouped_corrected_sizes_data) #displaying the corrected survey with corrected size data


```

#### -\>Plotting Bar Chart Of The Total Size In Each Bin(Eg:Small, Medium, Large, Extra Large)

```{r}
      #plotting bar chart of the total size in each bin
      ggplot(data = grouped_corrected_sizes_data,
       mapping = aes(x = corrected_sizes, 
                     y = count, 
                    fill = corrected_sizes)) + 
      geom_bar( stat="identity") +
      labs(title = "Number of survey participant by size of their group",
       subtitle = "Most common gruop size among participants was `small`",
       caption = "Source: Survey Dataset",
       x = "corrected sizes",
       y = "count")

```

### PART 4: RETRIEVING DATA OF PERCENTAGE EMPOYEES THAT ARE PHD STUDENTS, POSTDOCS AND PROFESSORS PER RESEARCH GROUP

```{r}

  group_perc_uni_data <-survey %>%
  group_by(university) %>%
  mutate( `percent_PhD_students` =(`number_of_phd_students`/`number_of_employees`*100),
          `percent_postDocs` =(`number_of_postdocs`/`number_of_employees`*100),
          `percent_professors` =(`number_of_professors`/`number_of_employees`*100))
  

  View(group_perc_uni_data)

```

#### -\> Reshaping Data For Stacked Plot Plotting Stacked Horizontal Bar Plot

```{r}

# Reshape data so that columns with percent are reshaped to a vertical shape
df_long <- group_perc_uni_data %>%
  pivot_longer(cols = starts_with("percent"), # Select columns to pivot
               names_to = "percent",          # New column for variable names
               values_to = "percentage_of_employees")        # New column for values


View(df_long) #displaying the percentage of employees survey for each research group id

# Creating the stacked horizontal bar plot
ggplot(df_long, 
        aes(x = `no`,y = `percentage_of_employees`, fill = `percent`)) +
        geom_col() +      
        coord_flip() +    
        labs(title = "Stacked Plot: Employees role percentage for each group",
             subtitle = "We can see that groups can have very different splits",
             y = "Percentage(%)",
             x = "Research group id",
             fill = "Employee Type") +
        theme_minimal() 



```

#### -\> Using Project1_Part4Chart Data (Provided in Resources Folder)

```{r}
project1_part4_chart_data <-  read_csv("data/project1_part4_chart_data.csv/project1_part4_chart_data.csv")

View(project1_part4_chart_data)

# Create the stacked horizontal bar plot
ggplot(project1_part4_chart_data, 
        aes(x = survey_id, y = percentage, fill = `employee_type`)) +
        geom_col() +      
        coord_flip() +  
        labs(title = "Stacked Horizontal Bar Plot ",
             subtitle = "Using project1_part4_chart_data",
             x = "Research group id",
             y = "Percentage",
             fill = "Employee Type") +
        theme_minimal() 

```

